On centrality of orthogonal $\Kt$

### 2016-05-11 paper.tex
Minor changes

It is a classical theorem of M.~Kervaire and R.~Steinberg that the stable Steinberg group $\St(R)$ is the universal central extension of $\E(R)$ for arbitrary $R$, see \cite[Theorem~III.5.5]{Kbook}.
The unstable analogue of the ``universality'' part of this theorem was obtained by M.~Stein and W.~van~der~Kallen for $\Phi$ of rank $\ell \geq 5$ (or even $\ell \geq 4$ for $\Phi$ of type $\rA_\ell, \rB_\ell, \rC_\ell$, see \cite{St1, KaS})
However, until recently, the analogue of the ``centrality'' part was only known in the case of a general $R$ for $\Phi=\rA_\ell$, $\ell\geq 3$ see~\cite[Corollary~2]{vdK}.
The best result for other classical $\Phi$ was that the centrality holds ``in the stable range'',
i.\,e. when the dimension of $R$ is suffiently small as compared to the rank of $\Phi$, see~\cite[Corollary~3.4]{St78}.

@@

We claim that \cref{tm:centrality} is a corollary of $\Kt$-analogue of Quillen's local-global principle,
very much in the same fashion as the normality theorem for the elementary subgroup is a corollary of Quillen's local-global principle for $\Ko$.
The proof of this claim can be found in \cite{SCh}
For the proof of this claim we refer the reader to the proof of \cite[Theorem~1]{SCh} which works in our situation verbatim.
%The above theorem is a corollary of the following local-global principle (cf. with the proof of~\cite[Theorem~1]{SCh}).
\begin{tm}[``Quillen's LG-principle for $\Kt$''] \label{tm:lg-principle} 
Let $\Phi$ be an irreducible simply-laced root system of rank $\geq 3$ and $R$ be arbitrary commutative ring.
Then an element $g\in\St(\Phi,\,R[X])$ satisfying $g(0)=1\in\St(\Phi,\,R)$ is trivial in $\St(\Phi,\,R[X])$
if and only if the images $\lambda^*(g)\in\St(\Phi,\,R_M[X])$ of $g$ under localisation maps $\lambda_M\colonR\to R_M$ are trivial for all maximal ideals $M\trianglelefteq R$. \end{tm}
The proof for the assertion that \cref{tm:lg-principle} follows from \cref{tm:centrality} is a verbatim repetion of the proof of \cite[Theorem~1]{SCh}.

\cref{tm:lg-principle} includes as a special case both~\cite[Theorem~2.1]{Tul} and~\cite[Theorem~2]{SCh} in which
the cases $\Phi=\rA_\ell$, $\ell\geq 4$, $\Phi=\rE_\ell$, $\ell=6,7,8$ were studied, respectively.
The special cases $\Phi=\rA_3$ and $\Phi=\rD_\ell$, $\ell\geq 4$ of \cref{tm:lg-principle} are new and have not been known before.

As a yet another motivation for ??? we mention that a symplectic counterpart of~\cref{lg-principle} was demonstrated by
the first-named author $\Phi=\rC_\ell$ starting from $\ell\geq 3$.

>>>>

It is a classical theorem of M.~Kervaire and R.~Steinberg that the stable Steinberg group $\St(R)$ is the universal central extension of $\E(R)$, see \cite[Theorem~III.5.5]{Kbook}.
The unstable analogue of the ``universality'' part of this theorem is known since 1970's from the works of M.~Stein and W.~van~der~Kallen for $\Phi$ with rank $\ell \geq 5$ (or even $\ell \geq 4$ for $\Phi=\rA_\ell, \rB_\ell, \rC_\ell$), see \cite{St1, KaS}.
However, until recently, the analogue of the ``centrality'' part was only known in the special case $\Phi=\rA_\ell$, $\ell\geq 3$ see~\cite[Corollary~2]{vdK}.
The best known result for other classical $\Phi$ was that the centrality holds ``in the stable range'',
i.\,e. when the dimension of $R$ is suffiently small as compared to the rank of $\Phi$, see~\cite[Corollary~3.4]{St78}.

@@

Centrality of $\Kt$ is a corollary of the following variant of Quillen's local-global principle,
very much in the same fashion as the normality theorem for the elementary subgroup is a corollary of Quillen's local-global principle for $\Ko$.
\begin{tm}[``Quillen's LG-principle for $\Kt$''] \label{tm:lg-principle} 
Let $\Phi$ be an irreducible simply-laced root system of rank $\geq 3$ and $R$ be arbitrary commutative ring.
Then an element $g\in\St(\Phi,\,R[X])$ satisfying $g(0)=1\in\St(\Phi,\,R)$ is trivial in $\St(\Phi,\,R[X])$
if and only if the images $\lambda_M^*(g)\in\St(\Phi,\,R_M[X])$ of $g$ under localisation maps $\lambda_M\colonR\to R_M$ are trivial for all maximal ideals $M\trianglelefteq R$. \end{tm}
For the proof of the claim that \cref{tm:lg-principle} follows from \cref{tm:centrality} we refer the reader to the proof of Theorem~1 in \cite{SCh} which works verbatim in our case.

The assertion of \cref{tm:lg-principle} was known previously in the following two situations:
\begin{itemize}
 \item in the case $\Phi=\rA_\ell$, $\ell\geq 4$ it was proved by Tulenbaev, the proof see~\cite[Theorem~2.1]{Tul}
 \item $\Phi=\rE_\ell$, $\ell=6,7,8$ and the second-named author
\end{itemize}
Thus, the special cases $\Phi=\rA_3$ and $\Phi=\rD_\ell$, $\ell\geq 4$ of \cref{tm:lg-principle} are new and have not been known before.
We also notice that a symplectic counterpart of~\cref{lg-principle} was demonstrated recently by the first-named author.

On the other hand, the special case 
includes as a special case both and~\cite[Theorem~2]{SCh} in which
the cases were studied, respectively.

As a yet another motivation for ??? we mention that
